---
author: Edwin Thoen
title: "A recipe for recipes"
output: ioslides_presentation
---

## The `recipes` package

"Preprocessing Tools to Create Design Matrices"

Define the steps you'll take to go from raw data to the analysis set.

Store these steps into a procedure (a recipe), and apply it on new data.

You focus on the best steps to take, `recipes` does the bookkeeping.

On CRAN since

## The `recipes` package

A `recipe` is the specification of an intent, separate the planning from the doing.

<picture of a recipe>

## Creating a `recipe`

```{r, warning = FALSE}
train_set <- mtcars[1:20, c("am", "disp", "hp")]
test_set  <- mtcars[21:32, c("am", "disp", "hp")]
library(recipes)
library(dplyr)
rec <- recipe(train_set, am ~ .)
```

We have defined the roles here, `am` is the outcome, `disp` and `hp` are the predictors.

## Adding `steps` to the `recipe`

Once a recipe has created we can add steps to it.

```{r}
rec_with_steps <- rec %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors())
```

This is specifying the intent, we didn't do anything on the data yet.

## Preparing the recipe with `prep`

Here the work starts. By applying `prep` to the recipe, we are acquiring all the necesarry information on the training set. 

```{r}
rec_prepped <- rec_with_steps %>% 
  prep(train = train_set)
rec_prepped
```

We now have the statistics to apply the centering and scaling derived from the training set and stored in the recipe.

## Preparing the data with `bake`

```{r}
train_final <- bake(rec_prepped, train_set)
test_final  <- bake(rec_prepped, test_set)
```

The statistics to center and scale are learned on the `train_set` and applied to the `test_set`.

## Why am I talking about `recipes`?

Added the `check` framework together with Max.

A `check` does not change the data in any way, it tests assumptions and will break `bake` if these are not met.

```{r}
rec <- recipe(train_set) %>% check_missing(everything()) %>% prep()
test_set[1, 1] <- NA
train_baked <- bake(rec, train_set)
test_baked  <- bake(rec, test_set)
```

## Building youw own steps and checks

Fully leverage package structure.

For your own preparations and to contribute to the package.

Challenge, delve a little deeper into the package inner workings.

## Aside: S3 objects

```{r}
eRum_generator <- function(name, country){
  x <- list(name = name, country = country)
  class(x) <- "eRum"
  x
}
eRum_generator("Edwin", "the Netherlands")

print.eRum <- function(x) {
  paste0(x$name, " from ", x$country, " is at eRum and thinks its awesome.") %>% 
    print()
}
eRum_generator("Edwin", "the Netherlands")
eRum_generator("Riccardo", "Italy")
```

## Aside: S3 objects

An S3 object is built by adding a `class` to a base object.

Functions that are specific for a class are called `methods`.

A `constructor` is a function that creates objects of a class in a uniform way.

## S3 classes in `recipes`

A recipe itself is of class `recipe`.

All the `steps` and `checks` available all have their own subclass. All with their own `prep` and `bake` functions. 

Recipe gathers all the objects of different subclasses.

When the methods `prep.recipe` and `bake.recipe` are called, they call the methods of the steps and checks.

## Create a custom `step` or `check`

A full step or check comprises:

- the function that is called to add to the recipe
- constructor to create new objects of the subclass
- `prep` method
- `bake` method
- `print` method
- `tidy` method

## A recipe for `recipes`

1) Write a function that does the baking action.
2) Recognize which arguments needs to be provided and which estimated by `prep`.
3) Write the constructor.
4) Write the function to be called.
5) Create the `prep` and `bake` methods.
6) Create the `print` and `tidy` methods. 

## Resources

Within `recipes` you'll find a number of helper functions.

Grab the source code from github.com/topepo/recipes to access them.

On github.com/EdwinTh/recipe_for_recipes/ you will find a skeleton.

## Example 1: A signed log

In *Practical Data Science with R* (Zumel and Mount) the authors define the signed log as:

if |x| < 1: 0
else:       sign(x) * log(|x|)

1) the function for this is:
```{r}
signed_log <- function(x, base = exp(1)) {
  ifelse(abs(x) < 1, 0, sign(x) * log(abs(x), base = base))
}
```

## Example 1: A signed log 

2) think about the arguments

We have one argument: `base`

Nothing has to be derived by `prep.check_log`

## Example 1: A signed log 

3) The constructor

```{r}

```

